First and foremost I would like to express my special appreciation and thanks to my advisor Prof. Karsten Schwan. It has been an honor to be his Ph.D. student. He has inspired me greatly to work in high performance computing (HPC) field. His willingness to motivate me and to help me grow as a researcher has contributed greatly to this thesis work. I am absolutely indebted to all his contributions of time, ideas, and research directions making my Ph.D. experience the most intellectually stimulating and enjoyable time in my career so far. I would also like to thank Dr. Matthew Wolf, Dr. Ada Gavrilovska and Dr. Jeffrey Young for mentoring and advising me at various points in my research. 

I would like to thank the other members of my dissertation committee, Prof. Sudhakar Yalamanchili, Prof. Ling Liu and Dr. Theodore L. Willke for serving as my committee members and for their insightful comments and suggestions on my research. My mentors during my internships at Intel Labs, Xia Zhu (Ivy) and Narayanan Sundaram, have had a major impact on shaping the ideas presented in this thesis. I would also like to express my special thanks to my manager at Intel Labs Theodore L. Willke, who is still a mentor to me. Susie McClain, our group admin, who has been extremely helpful in taking care of our travels, day to day lab needs, etc deserves a special approbation. Not many people can keep that amount of organization and detail in their head.

My time at Georgia Tech  was made enjoyable in large part due to many friends and groups that became a part of my life. I am grateful for the time spent and several intellectually stimulating discussions with my past and present labmates in the CERCS group over the years. A special mention to Minsung Jang, Ketan Bhardwaj, Bharath Srinivasan, Mukil Kesavan, Abhinav Narain, Sudarsun Kannan, Hrishikesh Amur, Vishal Gupta, Subramanya R. Dulloor, Min Lee, Fang Zheng, Hobin Yoon, Alex Merritt, Junwei Li and Jai Dayal. Mukil and Minsung were particularly very helpful in giving me general advice for a successful Ph.D. life. I hope to keep in touch with all of them in future. I am extremely grateful to Sudarsun for being the nicest roommate ever. I have learnt so much from him and he always has the most sound advice I can get on life issues. I would also like to thank all my friends from IIT Kharagpur: Subhra Mazumdar, Puneet Jain, Satya Gautam, Pranith Kumar, Karan Doshi and Anchal Nema for helping me feel lighthearted during the most stressful moments in my Ph.D. life. 

Finally, I would like to thank my family for their love and support, without whom I would have given up a long time ago. For my parents who have always supported me in all my pursuit be it my interest in higher studies or my love for liberal arts. For the presence of my elder brother one phone call away with his intellectually and spiritually stimulating advice. I would like to dedicate this thesis to my parents for what I am today is all because of them.	
